---
title: "PCA"
output: pdf_document
---


SITE: http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/118-principal-component-analysis-in-r-prcomp-vs-princomp/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(factoextra)
```




Iremos analisar a base de dados "meninas_unicas" obtidas do arquivo analise_estrutura.Rmd
```{r}
str(meninas_unicas)
```


```{r}
meninas_unicas$UF.de.Nascimento <- as.factor(meninas_unicas$UF.de.Nascimento)
meninas_unicas$Cotista <- as.factor(meninas_unicas$Cotista)
meninas_unicas$Tipo.de.Escola <- as.factor(meninas_unicas$Tipo.de.Escola)
meninas_unicas$Raça <- as.factor(meninas_unicas$Raça)
meninas_unicas$Curso <- as.factor(meninas_unicas$Curso)
meninas_unicas$Opção <- as.factor(meninas_unicas$Opção)
meninas_unicas$Forma.de.Ingresso.na.UnB <- as.factor(meninas_unicas$Forma.de.Ingresso.na.UnB)
meninas_unicas$Forma.de.saída.da.opção <- as.factor(meninas_unicas$Forma.de.saída.da.opção)
```




transformar os fatores em numeros
(DEVE SER NUMERICO)
```{r}
meninas_unicas$UF.de.Nascimento <- sapply(as.factor(meninas_unicas$UF.de.Nascimento), as.numeric)
meninas_unicas$Cotista <- sapply(as.factor(meninas_unicas$Cotista), as.numeric)
meninas_unicas$Tipo.de.Escola <- sapply(as.factor(meninas_unicas$Tipo.de.Escola), as.numeric)
meninas_unicas$Raça <- sapply(as.factor(meninas_unicas$Raça), as.numeric)
meninas_unicas$Curso <- sapply(as.factor(meninas_unicas$Curso), as.numeric)
meninas_unicas$Opção <- sapply(as.factor(meninas_unicas$Opção), as.numeric)
meninas_unicas$Forma.de.Ingresso.na.UnB <- sapply(as.factor(meninas_unicas$Forma.de.Ingresso.na.UnB), as.numeric)
meninas_unicas$Forma.de.saída.da.opção <- sapply(as.factor(meninas_unicas$Forma.de.saída.da.opção), as.numeric)
```


```{r}
str(meninas_unicas[, 4:13])
```



```{r}
#any(is.na(log.ir))
#sum( sapply(log.ir, which(is.infinite) )
#dim(log.ir)
```

The function princomp() uses the spectral decomposition approach. The functions prcomp() and PCA()[FactoMineR] use the singular value decomposition (SVD).

vizualizazr os eigenvalues

```{r}
res.pca <- prcomp(meninas_unicas[, 4:14], scale = TRUE)
fviz_eig(res.pca)
```

Graph of individuals. Individuals with a similar profile are grouped together

```{r}
fviz_pca_ind(res.pca,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = FALSE     # Avoid text overlapping
             )
```

Grafico das variáveis


```{r}
fviz_pca_var(res.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
```



```{r}
fviz_pca_biplot(res.pca, repel = FALSE,
                col.var = "#2E9FDF", # Variables color
                col.ind = "#696969"  # Individuals color
                )
```


PCA result

```{r}
library(factoextra)
# Eigenvalues
eig.val <- get_eigenvalue(res.pca)
eig.val
```


```{r}
# Results for Variables
res.var <- get_pca_var(res.pca)
res.var$coord          # Coordinates
res.var$contrib        # Contributions to the PCs
res.var$cos2           # Quality of representation 
# Results for individuals
res.ind <- get_pca_ind(res.pca)
res.ind$coord          # Coordinates
res.ind$contrib        # Contributions to the PCs
res.ind$cos2           # Quality of representation 
```


#Variáveis Categóricas
Qualitative / categorical variables can be used to color individuals by groups. The grouping variable should be of same length as the number of active individuals (here 23).

```{r}
groups <- as.factor(meninas_unicas_u$Forma.de.saída.da.opção)
fviz_pca_ind(res.pca,
             col.ind = groups, # color by groups
             addEllipses = TRUE, # Concentration ellipses
             ellipse.type = "confidence",
             legend.title = "Groups",
             repel = FALSE
             )
```

Calculate the coordinates for the levels of grouping variables. The coordinates for a given group is calculated as the mean coordinates of the individuals in the group.

```{r}
library(magrittr) # for pipe %>%
library(dplyr)   # everything else
# 1. Individual coordinates
res.ind <- get_pca_ind(res.pca)
# 2. Coordinate of groups
coord.groups <- res.ind$coord %>%
  as_data_frame() %>%
  select(Dim.1, Dim.2) %>%
  mutate(competition = groups) %>%
  group_by(competition) %>%
  summarise(
    Dim.1 = mean(Dim.1),
    Dim.2 = mean(Dim.2)
    )
coord.groups
```



```{r}
princomp(meninas_unicas[, 4:13], cor = FALSE, scores = TRUE)
```

